A Collection of Thoughts on the Keccak/SHA-3 Construction

### 2018-04-17 s2018-math490/thesis/thesis.tex
Updates to thesis
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

\subsection{Margin Properties} \label{sec:d:margin}

@@

@@

@@

>>>>

\subsection{Marginal Properties} \label{sec:d:margin}

@@

In this section, we introduce two potential attacks against Keccak/SHA-3 using
fixed points in the underlying round function. However, neither of these
attacks have been proven possible with current analysis except for small values
of $w$ and small numbers of rounds. The first is a possible attack using full
fixed points in the core round function, while the latter describes a category
of partial fixed points.

@@

One theoretical attack against SHA-3 is by using a fixed point. If there
existed a state $x$ such that $h(x) = x$, for $h$ the round function, then
for all additional blocks, $h(x | b) = h(x \oplus b)$. While fixed points can
and do occur (see 24-rounds in Table \ref{tab:p:c:1}), using a fixed-point
attack is more subtle in practice: it is unlikely that a fixed point is a valid
block (that is, $x$ contains a suffix that is $0^{m}$ for the current margin,
$m$). Thus one must find a series of blocks $b_1$, $b_2$ such that
$h(b_1) \oplus b_2 = x$; this in turn extends the attack from being
$h(x | b) = h(x \oplus b)$ to:
\begin{align*}
    x = h(h(b_1) \oplus b_2) & \Rightarrow \\
    h(x | b) = h(b_1 | b_2 | b) & = h(b_1 | h_2 \oplus b) \\
    & = h(b_1 | h_2 | 0^{25w} | b)
\end{align*}
However finding $b_1$, $b_2$ amounts to a preimage attack and is thus unlikely
to occur.

@@

Another theoretical attack is using a partial fixed point. Suppose there
exists a block $x$ such that $h(x) = y$ is also a block (that is, $y$ has
a suffix of $0^{m}$). In this case, the following is a valid collision,
for all blocks $b$:
\begin{align*}
    h(x \oplus b) = h(y | b)
\end{align*}
We have verified the existance of these partial fixed points for small
$w$ and number of rounds. Furthermore, they are more readily found using SAT
than finding a fixed point is.
